---
title: "Biostat776_Project_2"
author: 'Can Wang'
date: "2024-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1: Fun with functions
### Part 1A: Exponential transformation
```{r}
Exp <- function(x, k) {
    sum <- 1
    for(i in 1:k){
      sum <- sum + x^i/factorial(i)
    }
    return(sum)
}
Exp(5,100)
```

### Part 1B: Sample mean and sample standard deviation
```{r}

sample_mean <- function(x) {
  sum <- 0
  N <- length(x)
  for(i in 1:N){
    sum <- sum + x[i]
  }
  return(sum/N)
}

sample_sd <- function(x) {
    sum <- 0
    N <- length(x)
    for(i in 1:N){
      sum <- sum + (x[i]-sample_mean(x))^2
    }
    return(sqrt(sum/(N-1)))
}

x <- c(1,2,3,4,5,6,7,8,9,10)
print(sample_mean(x))
print(sample_sd(x))
```

### Part 1C: Confidence intervals
```{r}
calculate_CI <- function(x, conf = 0.95) {
    alpha <- 1 - conf
    N <- length(x)
    degrees_freedom <- N - 1
    t_score <- qt(p = alpha / 2, df = degrees_freedom, lower.tail = FALSE)
    lower_bound <- mean(x) - t_score*(sd(x)/sqrt(N))
    upper_bound <- mean(x) + t_score*(sd(x)/sqrt(N))
    CI <- c(lower_bound,upper_bound)
    return(CI)
}
calculate_CI(x,conf = 0.95)
calculate_CI(x,conf = 0.90)

## Compare with existing function
dat <- data.frame(x = x)
fit <- lm(x ~ 1, dat)
confint(fit, level = 0.95)
confint(fit, level = 0.90)
```

# Part 2: Wrangling data
### Data
```{r}
library(here)
if (!file.exists(here("data", "tuesdata_rainfall.RDS"))) {
    tuesdata <- tidytuesdayR::tt_load("2020-01-07")
    rainfall <- tuesdata$rainfall
    temperature <- tuesdata$temperature

    # save the files to RDS objects
    saveRDS(tuesdata$rainfall, file = here("data", "tuesdata_rainfall.RDS"))
    saveRDS(tuesdata$temperature, file = here("data", "tuesdata_temperature.RDS"))
}
rainfall <- readRDS(here("data", "tuesdata_rainfall.RDS"))
temperature <- readRDS(here("data", "tuesdata_temperature.RDS"))
library(tidyverse)
glimpse(rainfall)
glimpse(temperature)
```

### 1. Start with rainfall dataset and drop any rows with NAs.
```{r}
rainfall_omit <- na.omit(rainfall)
```

### 2. Create a new column titled date that combines the columns year, month, day into one column separated by “-”. (e.g. “2020-01-01”). This column should not be a character, but should be recognized as a date.
```{r}
rainfall_omit$date <- ymd(paste(rainfall_omit$year, rainfall_omit$month, rainfall_omit$day, sep = "-"))
```

### 3. Using the city_name column, convert the city names (character strings) to all upper case.
```{r}
rainfall_omit$city_name <- str_to_upper(rainfall_omit$city_name)
glimpse(rainfall_omit)
```

### 4. Join this wrangled rainfall dataset with the temperature dataset such that it includes only observations that are in both data frames. (Hint: there are two keys that you will need to join the two datasets together).
```{r}
df <- inner_join(rainfall_omit, temperature, by = c("city_name", "date"))
glimpse(df)
```

# Part 3: Data visualization
### Part 3A: Plotting temperature data over time
```{r}
df_3a <- df[c('city_name','year','temperature')]
df_3a <- df_3a[df_3a$year>=2014,]
df_3a_plot <- df_3a%>%
  group_by(city_name,year)%>%
  summarise(max=max(temperature,na.rm = TRUE),min=min(temperature,na.rm = TRUE),.groups = 'drop')


df_3a_plot_long <- df_3a_plot %>%
  pivot_longer(cols = c(max, min), 
               names_to = "temperature_type", 
               values_to = "temperature")

ggplot(df_3a_plot_long, aes(x = year, y = temperature, color = temperature_type)) +
  ggthemes::theme_stata()+
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Max and Min Temperatures Over Time",
       x = "Year",
       y = "Temperature",
       color = "Temperature Type") +
  facet_wrap(~ city_name) + 
  labs(subtitle="Key Trends: Maximum and minimum of temerature in the following cities remains stable over year. ",caption = "Created by Can Wang") 
```

### Part 3B: Plotting rainfall over time
```{r}
plot_rainfall <- function(p_city, p_year){
  filtered_data <- df %>%
    filter(city_name == p_city, year == p_year)
  
  if (nrow(filtered_data) == 0) {
    stop(paste("Error: No data found for city:", p_city, "and year:", p_year))
  }
  
  ggplot(filtered_data, aes(log(rainfall))) +
    ggthemes::theme_stata() +
    geom_histogram(fill = "skyblue") +
    labs(title = paste("Rainfall Distribution in", p_city, "in", p_year),caption="Created by Can Wang") 
}

plot_rainfall("SYDNEY", 2000)
```

Description: In the function plot_rainfall, there are two parameters, which are p_city (the city name) and p_year (the year). The user input the two parameters, and the function returns the histogram of log-rainfall of that particular city and year. In the function, we first filter the data by city_name and year. Then we check if the data exist. If not, use stop() to report this error. Otherwise, we then plot the histogram. The plot includes a title indicating the parameters, as well as informative x-axis and y-axis labels.

# Part 4: Apply functions and plot
### Part 4A: Apply the functions we wrote in Part 1 to our rainfall data starting with our wrangled df data from Part 2
```{r}
## First, filter for only years including 2014 and onwards.
df_4a <- df[df$year>=2014,]
## For a given city and for a given year, calculate the sample mean (using your function sample_mean()), the sample standard deviation (using your function sample_sd()), and a 95% confidence interval for the average rainfall (using your function calculate_CI()). Specifically, you should add two columns in this summarized dataset: a column titled lower_bound and a column titled upper_bound containing the lower and upper bounds for you CI that you calculated (using your function calculate_CI()).
rain_df <- df_4a %>%
  group_by(city_name,year)%>%
  summarise(mean_rainfall=sample_mean(rainfall),
            sd_rainfall=sample_sd(rainfall),
            lower_bound=calculate_CI(rainfall)[1],
            upper_bound=calculate_CI(rainfall)[2])
rain_df
```

### Part 4B: Using the rain_df, plots the estimates of mean rainfall and the 95% confidence intervals on the same plot. There should be a separate faceted plot for each city. 
```{r}
ggplot(rain_df, aes(x = year, y = mean_rainfall)) +
  ggthemes::theme_stata()+
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.2)+
  labs(title = "Mean rainfall and 95% CI",
       x = "Year",
       y = "Mean Rainfall") +
  facet_wrap(~ city_name) 
```



